Minimal trap spaces of Logical models are maximal siphons of their Petri net encoding

### 2022-05-06 cmsb22.tex
Details

Then, we translate the above characterization into the ASP \(\mathcal{A}\) as follows. We use atom \verb|p(v, 1)| (resp. \verb|p(v, 0)|) to denote place \(p_v\) (resp. \(\overline{p_v}\)), \(\forall v \in V\). For each pair \((p, t)\) where \(p \in P, t \in T, t \in pred(p)\), we translate the rule (\ref{eq:siphon}) into the ASP rule
\[
\verb|p(v_1, x_1); ... ; p(v_k, x_k) :- p(v, x).|
\]
where \verb|p(v, x)| is the atom representing place \(p\) and \{\verb|p(v_1, x_1)|, \dots,\linebreak \verb|p(v_k, x_k)|\} is the set of atoms representing places in \(pred(t)\). The rule (\ref{eq:conflict}) is translated into the ASP rule
\[\verb|:- p(v, 1), p(v, 0).|\]
for each \(v \in V\). This rule guarantees that \verb|p(v, 1)| and \verb|p(v, 0)| never belong to the same answer set of \(\mathcal{A}\), representing the conflict-freeness. Note that the number of atoms of \(\mathcal{A}\) is only \(2n\), which is more theoretically advanced than the ASP encoding shown in~\cite{klarner2015computing} where the number of atoms of the encoded ASP equals to the number of prime-implicants of the Boolean model and that number might be exponential in \(n\).

% \sylvain{What do you mean by \emph{advanced}? It is less since there might be an exponential number of prime-implicants?}
% \vangiang{Yes. Since there might be an exponential number of prime-implicants, our new ASP encoding is more advanced than the ASP encoding of PyBoolNet.}

@@

In this article we proposed a new method for the computation of minimal trap-spaces of Boolean models, based on a new concept called maximal conflict-free siphons.
This method is evaluated on big models from the literature and shows that it can scale up much better than the state-of-the-art prime-implicants based techniques.
We believe that this opens up the way to a much better analysis of large Boolean models, which is needed with the advent of automatic model-generation pipelines~\cite{ostaszewski2021covid19}.

>>>>

Then, we translate the above characterization into the ASP \(\mathcal{A}\) as follows. We use atom \verb|p(v, 1)| (resp. \verb|p(v, 0)|) to denote place \(p_v\) (resp. \(\overline{p_v}\)), \(\forall v \in V\). For each pair \((p, t)\) where \(p \in P, t \in T, t \in pred(p)\), we translate the rule (\ref{eq:siphon}) into the ASP rule
\[
\verb|p(v_1, x_1); ... ; p(v_k, x_k) :- p(v, x).|
\]
where \verb|p(v, x)| is the atom representing place \(p\) and \{\verb|p(v_1, x_1)|, \dots,\linebreak \verb|p(v_k, x_k)|\} is the set of atoms representing places in \(pred(t)\). The rule (\ref{eq:conflict}) is translated into the ASP rule
\[\verb|:- p(v, 1), p(v, 0).|\]
for each \(v \in V\). This rule guarantees that \verb|p(v, 1)| and \verb|p(v, 0)| never belong to the same answer set of \(\mathcal{A}\), representing the conflict-freeness. Note that the number of atoms of \(\mathcal{A}\) is only \(2n\), whereas the ASP encoding shown in~\cite{klarner2015computing} has as many atoms as the number of prime-implicants of the Boolean model and that number might be exponential in \(n\).

@@

In this article we proposed a new method for the computation of minimal trap-spaces of Boolean models, based on a new concept called maximal conflict-free siphons.
This method is evaluated on large models from the literature and shows that it can scale up much better than the state-of-the-art prime-implicants based techniques.
We believe that this opens up the way to a much better analysis of large Boolean models, which is needed with the advent of automatic model-generation pipelines~\cite{ostaszewski2021covid19}.
